Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-05-17 semantic-nd/article.tex
More small changes to everything before novelty detection section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
Multi-modal approaches, such as combining semantic data extracted from
several sources or classifiers are expected to have better performance on scene
recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
showed that most scene recognition models work poorly in indoor scenes when
compared to outdoor scenes since the properties that
characterize rooms changes depending on the category. For instance corridors are well
described by global properties and bookstores are well described by the presence of
specific objects (books).
Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
between object and room category, where object defines a room category and a room
category provides information on where objects may be found.

@@

Novelty detection has been studied for many years and there are several approaches
based on statistical analysis\cite{markou2003novelty}.
Graphical models have been used to learn distributions of variables, both in
supervised and unsupervised ways and by using thresholds on those distributions
based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
a novelty system can be trivially implemented.
However to the knowledge of the authors there is no reference on how to perform
novelty detection using graphs that are dynamically generated.

@@

\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}
\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}
\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
         in order to create a structured probabilistic representation of the world.}
\end{figure}

\subsection{Factor Graphs}
Although the conceptual layer works with \emph{chain graphs}, those can be converted
into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
easier manipulation due to factorization.
Moreover, there exists efficient implementation of inference engines operating on factor graph
representations\cite{Mooij_libDAI_10}.

Describing the distribution function in terms of graphs allows to use algorithms
such as the Sum-Product\cite{kschischang2001factor} to efficiently calculate
marginals on any given subset of variables by exploiting conditional
independence between variables.

A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
representing random variables and factors.
Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
to which the factor is connected.
Thus, a factor graph can be seen as a description of probability density function obtained
by a product of all the factors. In order to represent the probability,
a normalization factor needs to be introduced, resulting in the following equation
defining the distribution:

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
% TODO(pronobis): general comment about many .... approches to place categorization.

Multi-modal approaches, such as combining semantic data extracted from
several sources or classifiers are expected to have better performance on scene
recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
showed that most scene recognition models work poorly in indoor scenes when
compared to outdoor scenes since the properties that
characterize rooms changes depending on the category. For instance corridors are well
described by global properties and bookstores are well described by the presence of
specific objects (books).
Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
between object and room category, where object defines a room category and a room
category provides information on where objects may be found.

@@

Novelty detection has been studied for many years and there are several approaches
based on statistical analysis\cite{markou2003novelty}.
Graphical models have been used to learn distributions of variables, both in
supervised and unsupervised ways and by using thresholds on those distributions
based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
a novelty system can be trivially implemented.

However to the knowledge of the authors there is no reference on how to perform
novelty detection using graphs that are dynamically generated.
% TODO(andresp): anything more that can be said on why this paper is useful?

@@

\begin{figure}[h]
\centering
\subfloat[Conceptual layer]{\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}}
\subfloat[Chain graph]{\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}}

\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
         in order to create a structured probabilistic representation of the world.}
\end{figure}

\subsection{Factor Graphs}
Although the conceptual layer works with \emph{chain graphs}, those can be converted
into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
easier manipulation due to factorization.
Moreover, there exists efficient implementation of inference engines operating on factor graph
representations\cite{Mooij_libDAI_10}.
Describing the distribution function in terms of graphs allows to use those engines to
efficently calculate marginals on any given subset of variables by exploiting conditional
independence between variables.

A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
representing random variables and factors.
Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
to which the factor is connected.
Thus, a factor graph can be seen as a description of probability density function obtained
by a product of all the factors. In order to represent the probability,
a normalization factor needs to be introduced, resulting in the following equation:
